---
title: "Admixture analysis"
author: "Tina Lasisi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: true
    number_sections: yes
editor_options:
  chunk_output_type: console
---
```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(tidyverse)
library(knitr)
library(cowplot)
library(data.table)
library(modelr)
library(broom)

F = rprojroot::is_rstudio_project$make_fix_file()

knitr::opts_chunk$set(echo = FALSE, include = TRUE, eval = TRUE, warning = FALSE, message = FALSE, fig.retina = 2, fig.width = 8, fig.height = 4)

```

```{r functions, include=FALSE}
plot_path = F("output/")

pltsave_func <- function(plot, plot_path, width=6, height=7){
  ggsave(
      filename = paste0(deparse(substitute(plot)), ".png"),
      plot = plot, 
      path = F("output/"),
      width = width,
      height = height)
}

```


```{r read_df_pheno, include=FALSE}

# Import the phenotype data and wrangle into tidy format
df_pheno <- read_csv(F("data/ignore_data/df_Admixed_Merged.csv")) %>%
  group_by(ID, hair_texture, HairTypeBin, sex, age) %>%
  summarize(area_mean = mean(area),
         area_median = median(area),
         min_mean = mean(min),
         max_mean = mean(max),
         min_median = median(min),
         max_median = median(max),
         eccentricity_mean = mean(eccentricity),
         eccentricity_median = median(eccentricity),
         curv_median = mean(curv_median),
         curv_mean = mean(curv_mean),
         length_mean = mean(length_mean),
         length_median = mean(length_median),
         hair_count = mean(hair_count),
         m_index = mean(m_index))  %>%
  mutate(hair_texture = str_to_title(hair_texture)) %>% 
  ungroup()

df_pheno$hair_texture <- factor(df_pheno$hair_texture, 
                                     levels = c("Straight","Wavy",
                                                "Curly",
                                                "Very Curly",
                                                NA))

df_pheno$HairTypeBin <- factor(df_pheno$HairTypeBin, 
                                     levels = c("Straight","Wavy",
                                                "Curly",
                                                "Very Curly"))
```


```{r read_df_anc, include=FALSE}

# import the id key and the ancestry file, as needed, and wrangle to extract just the ancestry markers
# NB: when Jinguo gives you the new ancestry files, you may not have the same number of ancestry columns, so check

df_idkey <- read_csv(F("data/ignore_data/id_key.csv")) %>% 
  mutate(across(.fns = as.numeric))

df_anc <- read_csv(F("data/ignore_data/df_ancestry.csv")) %>% 
  rename(NID = ID) %>% 
  mutate(NID = as.numeric(NID))

df_anc <- left_join(x = df_anc, y=df_idkey) %>% 
  select(NID, ID, 18:22) %>% 
  unique()
  
colnames(df_anc)[c(3:7)] = c("SAS","AMR","EUR","AFR","EAS")
```


```{r df_merge, include=FALSE}

df_merged <- left_join(x = df_pheno, y = df_anc)

```

# Admixture components
Our sample consists of admixed individuals with primarily African and European ancestry.

We first calculated the ancestry for the entire ADAPT sample.
```{r plt_admixture_ADAPT, fig.height=5, fig.width=10, fig.cap= "Admixture components for hair project sample"}
# show admixture plot for entire ADAPT sample
```


Below we plot the ancestries for the admixed African-European sample that was phenotyped using the new hair quantification methods. 

```{r plt_admixture, fig.height=5, fig.width=10, fig.cap= "Admixture components for hair project sample"}

anc <- df_merged %>% 
  ungroup() %>% 
  select(ID, SAS, AMR, AFR, EUR, EAS) %>% 
  arrange(AFR)

anc$ID <- factor(anc$ID, levels = anc$ID)

manc = reshape2::melt(anc, id.vars="ID")

plt_manc = ggplot(manc)+
  geom_bar(aes(ID,value, fill=variable),
           stat = "identity",
           width=1)+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())+
  labs(x = "Individuals",
       y = "Ancestry",
       fill = "Population")+
  scale_fill_manual(values = c(
    "#e41a1c",
    "#377eb8",
    "#4daf4a",
    "#984ea3",
    "#ff7f00"
  ))

pltsave_func(plt_manc, height = 5)

```

The colors represent ancestries that correspond to the following 1000 Genomes populations:
- SAS = South Asian
- AMR = American
- AFR = African
- EUR = European
- EAS = East Asian

Each of these are metapopulations based on the grouping of multiple (sub)continental population groups in the [1000 Genomes repository](https://www.internationalgenome.org/category/population/).

# Ancestry vs. hair morphology 

We carried out a number of analyses using the genotype data collected for this diverse sample. In an admixed sample where a continuous trait has divergent distributions in the parental ancestry groups, the resulting admixed population can show a correlation between ancestry and that trait. Finding such a correlation suggests may imply a polygenic trait with high heritability.


## Ancestry vs. curvature

Here we plot the correlation between proportion of African ancestry and m-index, curvature, area, minimum diameter, maximum diameter, and area (using the median for each).


```{r df_residuals, include=FALSE}

# wrangle into long format and keep only the median values for the hair properties

df_merged_long <- df_merged %>% 
  select(ID, sex, age, AFR, m_index, contains("median") & !contains("length")) %>% 
  pivot_longer(cols = m_index:curv_median, names_to="phenotype")

df_resid <- df_merged_long %>% 
  group_by(phenotype) %>% 
  drop_na() %>% 
  nest() %>% 
  mutate(model = map(data, ~lm(value ~ AFR, data = .)),
         residuals = map2(data, model, add_residuals)) %>% 
  unnest(residuals) %>% 
  select(ID, sex, age, AFR, phenotype, value, resid)

df_resid_wide <- df_resid %>% 
  pivot_wider(names_from = phenotype, values_from = c(value, resid))

```

### Uncorrected

```{r plt_anc_pheno}


plt_anc_pheno<- ggstatsplot::grouped_ggscatterstats(
  data = df_merged_long,
  x = AFR,
  y = value,
  grouping.var = phenotype,
  bf.message = FALSE,
  marginal = FALSE,
  plotgrid.args = list(ncol = 2, labels = "auto"),
  ggstatsplot.layer = FALSE,
  ggplot.component = list(theme(text = element_text(size = 8)))
)

pltsave_func(plt_anc_pheno, height = 12, width = 10)

plt_corrmat <- ggstatsplot::ggcorrmat(
  data = df_resid_wide,
  cor.vars = contains("value")
)

pltsave_func(plt_corrmat)

```

### Corrected

# 1. skin vs. ancestry 
# 2. hair curvature, hair eccentricity, hair area 
# 3. ancestry corrected/residuals of all combinations pairwise
# lm between m-index and african ancestry, take residuals and test those residuals against the other phenotypes. Also m-index vs. those phenotypes without residuals. Then repeat every possible combination
# residualize each phenotype on ancestry then do a PCA


```{r plt_resid_anc_pheno}

plt_resid_anc_pheno<- ggstatsplot::grouped_ggscatterstats(
  data = df_resid,
  x = AFR,
  y = resid,
  grouping.var = phenotype,
  bf.message = FALSE,
  marginal = FALSE,
  plotgrid.args = list(ncol = 2, labels = "auto"),
  ggstatsplot.layer = FALSE,
  ggplot.component = list(theme(text = element_text(size = 8)))
)

pltsave_func(plt_resid_anc_pheno, height = 12, width = 10)

plt_resid_corrmat <- ggstatsplot::ggcorrmat(
  data = df_resid_wide,
  cor.vars = contains("resid")
)

pltsave_func(plt_resid_corrmat)

```




```{r imsave_mvc_evc_combo, include=FALSE}

plt_strat <- plot_grid(mvc_uncorrected, mvc_corrected, evc1, evc2, labels = c("a", "b", "c", "d"), nrow = 2, ncol = 2)

pltsave_func(plt_strat, plot_path, width = 8, height = 8)
```
