---
title: "Data QC"
author: "Tina Lasisi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: true
    number_sections: yes
editor_options:
  chunk_output_type: console
---


```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(tidyverse)
library(knitr)
library(cowplot)

F = rprojroot::is_rstudio_project$make_fix_file()

knitr::opts_chunk$set(echo = FALSE, include = TRUE, eval = TRUE, warning = FALSE, message = FALSE, fig.retina = 2, fig.width = 8, fig.height = 4, out.width = "100%")

```


```{r functions, include=FALSE}
plot_path = F("output/")

pltsave_func <- function(plot, plot_path, width, height){
  ggsave(
      filename = paste0(deparse(substitute(plot)), ".png"),
      plot = plot, 
      path = plot_path,
      width = width,
      height = height)
  plot(plot)
}

```


```{r random}

# images and other random analyses live in the old Hair Methods project folder
"/Users/tpl5158/old_2020_HairPheno_manuscript"

```


```{r read_df_admixed, include=FALSE}

df_admixed <- read_csv(F("data/ignore_data/renamed_df_Admixed_Merged.csv"))
colnames(df_admixed)[c(18:22)] = c("SAS","AMR","EUR","AFR","EAS")


```

To explore the significance of quantifying hair fiber morphology, we explore the relationship between various quantitative hair traits, categorical data and genotype data on the same sample. 


```{r df_admixed_plt, include=FALSE}
df_admixed_plt <- df_admixed %>%
  group_by(ID, hair_texture, HairTypeBin, sex, age) %>%
  summarize(area_mean = mean(area),
         area_median = median(area),
         min_mean = mean(min),
         max_mean = mean(max),
         min_median = median(min),
         max_median = median(max),
         eccentricity_mean = mean(eccentricity),
         eccentricity_median = median(eccentricity),
         curv_median = mean(curv_median),
         curv_mean = mean(curv_mean),
         length_mean = mean(length_mean),
         length_median = mean(length_median),
         SAS = mean(SAS),
         AMR = mean(AMR),
         EUR = mean(EUR),
         AFR = mean(AFR),
         EAS = mean(EAS),
         hair_count = mean(hair_count),
         m_index = mean(m_index))  %>%
  mutate(hair_texture = str_to_title(hair_texture)) 
# %>% 
#   filter(hair_count > 4 & AFR > 0.1)

df_admixed_plt$hair_texture <- factor(df_admixed_plt$hair_texture, 
                                     levels = c("Straight","Wavy",
                                                "Curly",
                                                "Very Curly",
                                                NA))

df_admixed_plt$HairTypeBin <- factor(df_admixed_plt$HairTypeBin, 
                                     levels = c("Straight","Wavy",
                                                "Curly",
                                                "Very Curly"))

# I filtered out anyone who had less than 5 hair fragments and I only kept the "first" reported ethnicity to make sure I didn't have duplicates of IDs
```

# Missing data

```{r df_na_admixed, include=FALSE}


# df_admixed_plt %>% 
#   print(is.na)

rowAny <- function(x) rowSums(x) > 0 
     
df_admixed_na <- df_admixed %>% 
  filter(rowAny(
         across(
             .cols = area:max,
             .fns = ~ is.na(.x)
         )
     )
)

write.csv(df_admixed_na, file = "data/ignore_data/admixed_na.csv", row.names = FALSE)

```

## Missing images

Below we are going to extract all the images for individuals with missing cross-sections

```{r}
# extracting new id numbers form df
list_id_missing <-df_admixed_na$ID 

pattern_id_missing <- paste(list_id_missing, sep = "", collapse = "|")

# image location path
orig_im_path <- "/Users/tpl5158/Box/01_TPL5158/Box_Dissertation/Repurpose/AfricanAmerican_sections_raw_copies/"

bramel_path <- "/Users/tpl5158/Box/01_TPL5158/Bramel_collab/data/raw/Raw_images_renamed"

AfrEu_path <- "/Users/tpl5158/Box/01_TPL5158/Box_Dissertation/Repurpose/HairPhenotyping_Methods/data/fibermorph_input/admixed_real_hair/section/AfrEu_SectionImages_RawJPG"

# new folder
dir_qc_missing <- "/Users/tpl5158/Box/01_TPL5158/Box_Dissertation/Afr_Admix_Assoc/data/ignore_data/qc/missing"


# list images from vector
list_id_missing <- list.files(path = AfrEu_path, pattern = pattern_id_missing, full.names = TRUE, recursive = TRUE) %>% 
  str_subset("^.*\\.RW2")

file.copy(list_id_missing, dir_qc_missing)

# list.files(path = AfrEu_path, pattern = pattern_nid_missing, full.names = TRUE, recursive = TRUE)





```

## Images with issues

```{r}

id_key <- F("data/ignore_data/id_key.csv") %>% 
  read_csv() %>% 
  mutate(ID = as.character(ID))
# import csv bad_im
list_bad_im <- F("data/ignore_data/bad_im.csv") %>% 
  read_csv(col_names = FALSE)

# import csv flawed_im
list_flawed_im <- F("data/ignore_data/flawed_im.csv")%>% 
  read_csv(col_names = FALSE)

# merge
list_issues_im <- bind_rows(list_flawed_im, list_bad_im) %>% 
  unique()

# rename to NID
fn_nid_issues_df <- list_issues_im %>% 
  separate(col = X1, into = c("ID", "rest"), sep = "_", extra = "merge") %>% 
  left_join(., id_key) %>% 
  # drop NA for now, but there are hairs that only were scanned in the first round that didn't make it through the new methods
  drop_na() %>% 
  select(NID, rest) %>% 
  unite("fn", NID:rest)

fn_nid_issues_list <- fn_nid_issues_df$fn

fn_id_issues_df <- list_issues_im %>% 
  separate(col = X1, into = c("ID", "rest"), sep = "_", extra = "merge") %>% 
  left_join(., id_key) %>% 
  # drop NA for now, but there are hairs that only were scanned in the first round that didn't make it through the new methods
  drop_na() %>% 
  select(ID, rest) %>% 
  unite("fn", ID:rest)

fn_id_issues_list <- fn_id_issues_df$fn

# new folder
dir_qc_issues <- "/Users/tpl5158/Box/01_TPL5158/Box_Dissertation/Afr_Admix_Assoc/data/ignore_data/qc/issues"

# find nid
pattern_nid_qc <- paste(fn_nid_issues_list, sep = "", collapse = "|")

# find by id
pattern_id_qc <- paste(fn_id_issues_list, sep = "", collapse = "|")

list_qc_issues <- list.files(path = AfrEu_path, pattern = pattern_id_qc, full.names = TRUE, recursive = TRUE)

file.copy(list_qc_issues, dir_qc_issues)



```

## Merge missing and issues image lists

```{r}

# merge lists together into final list
id_qc <- c(fn_id_issues_list, list_id_missing) %>% 
  unique()

# copy all images into a general fixing folder
dir_qc_manual <- F("data/ignore_data/im_manual_fix")

pattern_nid_qc <- paste(nid_qc, sep = "", collapse = "|")

list_qc_fn <- list.files(path = orig_im_path, pattern = pattern_nid_qc, full.names = TRUE)

file.copy(list_qc_fn, dir_qc_manual)

```


# Corrupted images
It turns out that some of the images were corrupted, so I made a subset that was just the corrupted images so I could search for the originals. 

```{r}

df_corr_im <- F("data/ignore_data/corrupted_im.csv") %>% 
  read_csv()

list_nid_corrupt <-df_corr_im$ID

pattern_nid_corr <- paste(list_nid_corrupt, sep = "", collapse = "|")

# image location path
orig_im_path <- "/Users/tpl5158/Box/01_TPL5158/Box_Dissertation/Repurpose/AfricanAmerican_sections_raw_copies/"

# list images from vector
fn_nid_corr <- list.files(path = orig_im_path, pattern = pattern_nid_corr, full.names = TRUE, recursive = TRUE) %>% 
  str_subset("^.*\\.RW2")

# new folder
dir_qc_corr <- F("data/ignore_data/qc/corrupt_src")

file.copy(fn_nid_corr, dir_qc_corr)

```


Once the images have been reconverted from raw to tiff, I'm going to rename them according to the key

```{r}

section_files <- tibble(list.files(F("data/ignore_data/qc/corrupt_src_tiff/tiff")))
colnames(section_files)[1] = "old_fn"

section_files <- section_files %>% 
  mutate(ID = as.character(old_fn)) %>% 
  filter(!str_detect(ID, "P")) %>% 
  separate(col = "ID", into = c("ID", "side", "hair", NA))

rename_section_df <- left_join(section_files, id_key) %>% 
  unite("new_fn", c("NID", "side", "hair")) %>% 
  mutate(new_fn = paste0(new_fn, ".tiff")) %>% 
  select(-ID)

```


```{r rename_section}
setwd(F("data/ignore_data/qc/corrupt_src_tiff/tiff"))

file.rename(from = as.vector(rename_section_df$old_fn), to = as.vector(rename_section_df$new_fn))

```

### Repeat for missing images
Once the images have been reconverted from raw to tiff, I'm going to rename them according to the key

```{r}

section_files <- tibble(list.files(F("data/ignore_data/qc/missing")))
colnames(section_files)[1] = "old_fn"

section_files <- section_files %>% 
  mutate(ID = as.character(old_fn)) %>% 
  filter(!str_detect(ID, "P")) %>% 
  separate(col = "ID", into = c("NID", "side", "hair", NA))

rename_section_df <- left_join(section_files, id_key) %>% 
  unite("new_fn", c("ID", "side", "hair")) %>% 
  mutate(new_fn = paste0(new_fn, ".tiff")) %>% 
  select(-NID)

pattern_nid_corr <- paste(list_nid_corrupt, sep = "", collapse = "|")

# image location path
orig_im_path <- "/Users/tpl5158/Box/01_TPL5158/Box_Dissertation/Repurpose/AfricanAmerican_sections_raw_copies/"

# list images from vector
fn_nid_corr <- list.files(path = orig_im_path, pattern = pattern_nid_corr, full.names = TRUE, recursive = TRUE) %>% 
  str_subset("^.*\\.RW2")

# new folder
dir_qc_corr <- F("data/ignore_data/qc/corrupt_src")

file.copy(fn_nid_corr, dir_qc_corr)

```


```{r rename_section2}
setwd(F("data/ignore_data/qc/missing"))

file.rename(from = as.vector(rename_section_df$old_fn), to = as.vector(rename_section_df$new_fn))

setwd(F("."))

```

# New merge
After re-running the data, we can remerge all the information - curvature, cross-section, ancestry and PEQ

```{r}

# import cross-sectional results
new_sect <- read_csv(F("data/ignore_data/qc/section_reanalyzed/Feb18_1604_fibermorph_section/summary_section_data.csv")) %>% 
  separate(col = "ID", into = c("ID", "side", "hair", NA)) %>% 
  rename(NID = ID) %>% 
  left_join(x=., y=id_key)

new_sect2 <- read_csv(F("data/ignore_data/qc/section_reanalyzed/Feb18_1842_fibermorph_section/summary_section_data.csv")) %>% 
  separate(col = "ID", into = c("ID", "side", "hair", NA)) %>% 
  left_join(x=., y=id_key)

old_sect <- read_csv(F("data/ignore_data/qc/section_reanalyzed/old_summary_section_data.csv"))%>% 
  separate(col = "ID", into = c("ID", "side", "hair", NA)) 

  
df_AdmixedSection <- bind_rows(old_sect, new_sect, new_sect2) %>% 
  select(-NID)

# This CSV file contains the latest curvature measurement variable for the Afr-Eur admixed sample.
df_AdmixedCurv <- dplyr::as_tibble(read_csv(F("data/ignore_data/curvature_summary_data_Nov02_1334.csv"))) %>% 
  select(-X1) %>% 
  separate(col = ID, into = c(NA, "ID"), sep = "-|_")

# This file contains m-index
AfAd_mergedMar24 <- as_tibble(read_csv(F("data/ignore_data/AfAd_mergedMar24.csv"))) %>% 
  select(ID, m_index) %>% 
  mutate(ID = as.character(ID))

# This CSV file contains the self-reported "group" the participants indicated and their self-reported hair texture if they answered that question on the Pre-Enrollment Questionnaire
df_ADAPT_SelfRep <- dplyr::as_tibble(read_csv(F("data/ignore_data/selfRepAncestryHair_complete_ADAPT.csv")))%>% 
  rename(
    ID = 'Study ID',
    age = Age,
    SelfRepGroup = Group,
    SelfRepComment = 'Clarify?',
    participant_SelfRep = Participant,
    father_SelfRep = Father,
    mother_SelfRep = Mother,
    fathers_mother_SelfRep = "Father's Mother",
    fathers_father_SelfRep = "Father's Father",
    mothers_mother_SelfRep = "Mother's Mother",
    mothers_father_SelfRep = "Mother's Father",
    hair_texture = "Hair Texture",
    important_ancestry_features = "Important ancestry features"
  ) %>% 
  mutate(
    ID = as.character(ID)
  ) %>% 
  select(-starts_with("X")) %>% 
  select(ID, age, participant_SelfRep, hair_texture)

library(readxl)

df_ParticipantPlinkInfo <- as_tibble(read_excel(F("data/ignore_data/Participants_JW.xlsx"), col_types = c("numeric", "text", "text","text", "numeric", "numeric", "text","text", "text", "text", "text"))) %>% 
  mutate(ID = as.character(ID),
         age = as.numeric(Age),
         sex = as.character(Sex)) %>% 
  mutate(sex = factor(case_when(
    sex == "1" ~ "male",
    sex == "2" ~ "female"
  ))) %>% 
  select(ID, sex)

```


Now merging everything together

```{r}

df_Admixed_Merged <- df_AdmixedCurv %>% 
  left_join(df_AdmixedSection, by = "ID") %>% 
  left_join(df_ADAPT_SelfRep, by = "ID") %>% 
  left_join(df_ParticipantPlinkInfo, by = "ID") %>%
  left_join(AfAd_mergedMar24, by = "ID") %>% 
  select(ID, age, sex, participant_SelfRep, hair_texture, everything()) %>% 
  mutate(
    participant_SelfRep = as_factor(participant_SelfRep),
    hair_texture = fct_relevel(as_factor(hair_texture), c("straight", "wavy", "curly", "very curly")),
    HairTypeBin = 
           factor(case_when(
             curv_median<= (1/54) ~ 'Straight',
             curv_median<= (1/28.5) ~ 'Wavy',
             curv_median<= (1/6) ~ 'Curly',
             TRUE ~ 'Very Curly'), levels = c("Straight", "Wavy", "Curly", "Very Curly")),
         HairShapeBin = 
           factor(case_when(
             eccentricity <= 0.6 ~ 'round',
             TRUE ~ 'elliptical'), levels = c("round", "elliptical"))
  )
```

```{r}
# find missing

df_admixed_na2 <- df_Admixed_Merged %>% 
  filter(rowAny(
         across(
             .cols = area:max,
             .fns = ~ is.na(.x)
         )
     )
)

write_csv(df_Admixed_Merged, file = "data/ignore_data/df_Admixed_Merged_20210218.csv")

```


Our data consists of 193 individuals for whom we have quantitative hair phenotype data. In our first data quality control step, we filter to keep individuals who have more than 4 hair fragments in their curvature image and over 10% African ancestry. We calculate mean and median values for the cross-sectional data we have collected for individuals (~ 6 sectioned hair fibers). In our analyses, we use median values as they are less affected by intra-individual outliers.


# Self-reported hair texture vs. quantitative hair curvature

We compare the self-reported hair texture with mean and median curvature for our sample. 

```{r plt_admixed_SelfRepHair, fig.height=4, fig.width=8, fig.cap= "Self-reported hair texture vs. quantitative hair curvature"}


plt_admixed_SelfRepHair_CurvMean <- df_admixed_plt %>%
  filter(is.na(hair_texture)=="FALSE")%>%
  ggplot(., aes(hair_texture,
                      curv_mean))+
  geom_violin(fill="#a6bddb",
              scale = "width")+
  geom_point(position = position_jitter(width=0.1),
             alpha = 0.4)+
  theme(axis.text.x = element_text(angle=90))+
  labs(x = "Self reported hair texture",
       y = bquote("Mean curvature ("*mm^-1*")"),
       title = "Quantitative variation underlying \n self reported hair texture (mean)")+
  theme_bw() +
  coord_flip()


plt_admixed_SelfRepHair_CurvMedian <- df_admixed_plt %>%
  filter(is.na(hair_texture)=="FALSE")%>%
  ggplot(., aes(hair_texture,
                      curv_median))+
  geom_violin(fill="#a6bddb",
              scale = "width")+
  geom_point(position = position_jitter(width=0.1),
             alpha = 0.4)+
  theme(axis.text.x = element_text(angle=90))+
  labs(x = "Self reported hair texture",
       y = bquote("Median curvature ("*mm^-1*")"),
       title = "Quantitative variation underlying \n self reported hair texture (median)")+
  theme_bw() +
  coord_flip()

plt_SelfRepCurv <- plot_grid(plt_admixed_SelfRepHair_CurvMean,
          plt_admixed_SelfRepHair_CurvMedian, labels = c("a", "b"))

plt_SelfRepCurv

```
